### 2021-03-05 Untitled.Rmd
post_favor_Sanders

---
title: "Untitled"
output: pdf_document
---

## post favor by treat
```{r postfavor, fig.cap="Distribution of feelings towards Sanders in six different situations",echo=FALSE}
#histogram for each treat
EDA<-EDA%>%
  drop_na(post_favor_Sanders)

ggplot(EDA, aes(x = post_favor_Sanders,fill=treat)) +
  geom_histogram(bins=15,color="white") +
  facet_grid(treat ~ .)+
  scale_fill_manual(values = c("#66C2A5", "#FC8D62", "#8DA0CB", "#E78AC3", "#A6D854", "#FFD92F"))+
  xlab("Feelings towards Sanders")+
  ggtitle("Distribution of feelings towards Sanders in six different situations")+
  theme_minimal()
# Similar distribution

#bar chart (table?)
post_S<-EDA%>%
  group_by(treat)%>%
  summarise(Average_mark=mean(post_favor_Sanders))

ggplot(aes(x=treat,y=Average_mark),data=post_S)+
  geom_bar(stat="identity",fill=c("#66C2A5", "#FC8D62", "#8DA0CB", "#E78AC3", "#A6D854", "#FFD92F"))+
  ggtitle("Distribution of feelings towards Sanders in six different situations")+
  xlab("Feelings towards Sanders")

#Takeaways: Compared with the control group, the video format with a scandal will increase negative affect toward Sanders.
```

``
# Average deception level by treat (result)
```{r table1,tab.cap="Average deception level of each media format", echo = FALSE}

EDA%>%
  group_by(treat)%>%
  filter(treat != "control")%>%
  summarise(Average=mean(believed_true,na.rm = T))%>%
  arrange(desc(Average))%>%
  knitr::kable(caption = "Average deception level of each media format",
               col.names = c("Media format", "Average Deception Level"))

#Takeaways:
# The video format was not better than audio or text in deceiving under presenting the same information.

```